\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{fancyvrb}
\usepackage{enumitem}
\setlist{nosep}

\title{Chat Application Deployment Guide}
\author{Course Project}
\date{\today}

\begin{document}

\maketitle

\section*{Overview}

This document explains how to deploy the Chat Application locally.
It describes all required dependencies, database setup, build steps,
and how to verify that the system is working correctly.

The repository also includes a working script called \texttt{deploy.sh}
that automates most of the setup process.

\section*{Project Structure}

The repository contains the following important directories:

\begin{verbatim}
client/
business/
data/
service/
sql/
Makefile
deploy.sh
README.md
\end{verbatim}

Each layer corresponds to the layered architecture required by the project.

\begin{itemize}
\item Client layer: HTML and test client
\item Service layer: HTTP server
\item Business layer: application logic
\item Data layer: database access
\item SQL directory: database schema and seed data
\end{itemize}

\section*{Required Software}

The following dependencies must be installed before building the project.

\subsection*{Arch Linux}

\begin{verbatim}
sudo pacman -Syu
sudo pacman -S base-devel mariadb gcc make asio
\end{verbatim}

\subsection*{Debian / Ubuntu}

\begin{verbatim}
sudo apt update
sudo apt install build-essential mariadb-server \
                 default-libmysqlclient-dev libasio-dev
\end{verbatim}

Required components:

\begin{itemize}
\item MariaDB database server
\item MariaDB client development libraries
\item ASIO networking library
\item g++ compiler with C++17 support
\item make build tool
\end{itemize}

Crow headers are already included in the repository.


\textbf{!!! ATTENTION !!!!}
To do eveything below, you can simply run the included deploy.sh script
by running:

\begin{verbatim}
./deploy.sh
\end{verbatim}



\section*{Database Setup}
Start the MariaDB server:

\begin{verbatim}
sudo systemctl start mariadb
sudo systemctl enable mariadb
\end{verbatim}

Create the database and user:

\begin{verbatim}
sudo mariadb <<SQL
CREATE DATABASE IF NOT EXISTS chat_app;
CREATE USER IF NOT EXISTS 'chat_user'@'localhost'
IDENTIFIED BY 'chatpass';
GRANT ALL PRIVILEGES ON chat_app.* TO 'chat_user'@'localhost';
FLUSH PRIVILEGES;
SQL
\end{verbatim}

Import the schema and seed data:

\begin{verbatim}
mariadb -u chat_user -pchatpass chat_app < sql/tables.sql
mariadb -u chat_user -pchatpass chat_app < sql/insert.sql
\end{verbatim}

\section*{Building the Project}

From the root of the repository run:

\begin{verbatim}
make clean
make
\end{verbatim}

This compiles the server binary.

\section*{Running the Server}

Start the server from the repository root directory.

\begin{verbatim}
./mychat-server
\end{verbatim}

Then open the application in a web browser:

\begin{verbatim}
http://localhost:18080
\end{verbatim}

\section*{Verification}

To confirm the deployment was successful:

\begin{enumerate}
\item Open the web interface in a browser.
\item Create a user through the UI.
\item Verify the user exists in the database.

\begin{verbatim}
mariadb -u chat_user -pchatpass chat_app -e \
"SELECT * FROM users;"
\end{verbatim}

\item Create a conversation through the UI.
\item Send a message.
\item Verify the message in the database.

\begin{verbatim}
mariadb -u chat_user -pchatpass chat_app -e \
"SELECT * FROM messages;"
\end{verbatim}

\item Edit a message in the UI.
\item Confirm the message text changed in the database.
\item Delete a message and confirm it is removed.
\end{enumerate}

\section*{Testing the API}

Example API test using curl:

\begin{verbatim}
curl http://localhost:18080/users
\end{verbatim}

Example SQL query for recent messages:

\begin{verbatim}
mariadb -u chat_user -pchatpass chat_app -e \
"SELECT * FROM messages ORDER BY message_id DESC LIMIT 10;"
\end{verbatim}

\section*{Troubleshooting}

Common problems:

\begin{itemize}
\item Database does not exist

Create it using:

\begin{verbatim}
CREATE DATABASE chat_app;
\end{verbatim}

\item Server returns 404 on the root page

Ensure the server is started from the repository root directory.

\item MariaDB started with skip grant tables

Restart MariaDB normally:

\begin{verbatim}
sudo systemctl restart mariadb
\end{verbatim}

\end{itemize}

\section*{Screenshot Checklist}

The following screenshots should be included in the final report.

\begin{enumerate}
\item Web interface loaded in browser
\item Creating a user
\item Database showing inserted user
\item Creating a conversation
\item Sending a message
\item Editing a message
\item Deleting a message
\item Running the automated test client
\end{enumerate}

\end{document}
